---
title: "Week2"
author: "Yue Jiang"
date: "2019/7/6"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1.Loading and preprocessing the data
```{r, echo=TRUE}
activity <- read.table("D:/dataset/activity.csv", sep=",", header=T)
head(activity)
```

2.What is mean total number of steps taken per day?
```{r echo=TRUE}
NA_index <- is.na(as.character(activity$steps))
narm_data <- activity[!NA_index,]
totalSteps <- aggregate(steps ~ date, data = activity, sum, na.rm = TRUE)
daily_steps <- aggregate(steps ~ date, data = narm_data, sum)
colnames(daily_steps) <- c("date", "steps")
head(daily_steps)
```


(2)If you do not understand the difference between a histogram and a barplot, research the difference between them. Make a histogram of the total number of steps taken each day.
```{r echo=TRUE}
hist(daily_steps$steps, breaks = 20, col = "blue", xlab = "Steps", main= "Total number of steps each day")
```

(3)Calculate and report the mean and median of the total number of steps taken per day

```{r echo=TRUE}
mean(daily_steps$steps)
median(daily_steps$steps)
```

3.What is the average daily activity pattern?

(1)Make a time series plot (i.e. \color{red}{\verb|type = "l"|}type="l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r echo=TRUE}
mean_steps <- mean(totalSteps$steps)
median_steps <- median(totalSteps$steps)
steps_interval <- aggregate(steps ~ interval, data = activity, mean, na.rm = TRUE)
plot(steps ~ interval, data = steps_interval, type = "l", xlab = "Time Intervals (5-minute)", ylab = "Mean number of steps taken (all Days)", main = "Average number of steps Taken at 5 minute Intervals",  col = "blue")
```

(2)Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r echo=TRUE}
maxStepInterval <- steps_interval[which.max(steps_interval$steps),"interval"]
maxStepInterval
```
4.Imputing missing values
(1)Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with \color{red}{\verb|NA|}NAs)
```{r echo=TRUE}
missing_rows <- sum(!complete.cases(activity))
missing_rows
```
(2)Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.
```{r echo=TRUE}
## This function returns the mean steps for a given interval
getMeanStepsPerInterval <- function(interval){
    steps_interval[steps_interval$interval==interval,"steps"]
}
```

(3)Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r echo=TRUE}
complete.activity <- activity
flag = 0
for (i in 1:nrow(complete.activity)) {
    if (is.na(complete.activity[i,"steps"])) {
        complete.activity[i,"steps"] <- getMeanStepsPerInterval(complete.activity[i,"interval"])
        flag = flag + 1
        }
}
```

(4)Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

```{r echo=TRUE}
total.steps.per.days <- aggregate(steps ~ date, data = complete.activity, sum)
hist(total.steps.per.days$steps, breaks=20, col = "blue", xlab = "Total Number of Steps", 
     ylab = "Frequency", main = "Histogram of Total Number of Steps taken each Day")
showMean <- mean(total.steps.per.days$steps)
showMedian <- median(total.steps.per.days$steps)
```
The mean of the complete dataset (10766) is equal to the mean of the dataset without missing values while the median shifted from 10765 to 10766. Therefore, the mean and median are almost identical.

5.Are there differences in activity patterns between weekdays and weekends?
(1)Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day
```{r echo=TRUE}
complete.activity$day <- ifelse(as.POSIXlt(as.Date(complete.activity$date))$wday%%6 == 
                                    0, "weekend", "weekday")
complete.activity$day <- factor(complete.activity$day, levels = c("weekday", "weekend"))
```


(2)Make a panel plot containing a time series plot (i.e. \color{red}{\verb|type = "l"|}type="l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.

```{r echo=TRUE}
steps.interval= aggregate(steps ~ interval + day, complete.activity, mean)
library(lattice)
xyplot(steps ~ interval | factor(day), data = steps.interval, aspect = 1/2, 
       type = "l")
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
